---
title: "Hw04-gapminder"
output: 
  html_document:
    keep_md: true
---

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(reshape2))
```

## Data Reshaping Prompts
### Activty #2

Here I'd like to make a tilble for yearly life expectancy for China, Japan and Canada, with the spread() function:
```{r}
lifeExp_sub <- gapminder %>%
  select("lifeExp","country","year") %>% 
  filter(country == "China" | country == "Canada" | country == "Japan")
  
  # using the spread() function to list the lifeExp with one row per year in three countries
  spread(lifeExp_sub,key = "country", value = "lifeExp") %>% 
  rename(lifeExp_Canada = "Canada", lifeExp_China = "China", lifeExp_Japan = "Japan") %>% 
  # stylize and show the table 
  knitr::kable(format = "html", caption = "lifeExp in three countries")
```

And plot the table above to show the lifeExp change in one graph:
```{r}
  ggplot(lifeExp_sub, aes(year,lifeExp)) +
  geom_point(aes(colour=country))
```

### Activty #5 Simple data manipulations in R
#### Group
First select the columns to form a table I want to work on:
```{r}
gdpPCap <- gapminder %>%
  select("gdpPercap","continent","country","year")
knitr::kable(head(gdpPCap,20))
```

Then we can group it by year and continent, and take the mean of each year's gdpPercap:
```{r}
gdp_group <- gdpPCap %>% 
  group_by(year,continent) %>% 
  summarize(mean_gdpPercap=mean(gdpPercap),nCountries=n())
knitr::kable(head(gdp_group,20))
```

#### SPLIT
```{r}
gdp_split <- gdp_group %>% 
 spread(key = "year", value="mean_gdpPercap")
knitr::kable(gdp_split)
```

#### STACK
Or stack the table above, with a union column of continent and number of countries:
```{r}
gdp_stack <- gdp_split %>% 
  unite(continent_n, continent, nCountries) %>% 
  melt(id="continent_n") %>% 
  arrange(continent_n)
knitr::kable(head(gdp_stack,30))
```

#### JOIN
```{r}
A<-data.frame(c1=c("C","a","n","a","d","a"), c2=1:6)
B <- data.frame(c1=c("C","h","i","n","a"), c3=c("o","p","q","r","s"))
C <- left_join(A, B)
A
B
knitr::kable(C)
```

#### SUBSET
Let's make a subset of 'gdp_group': filter the values in 2007.
```{r}
gdp_2007 <- gdp_group %>% 
  filter(year == 2007)
knitr::kable(gdp_2007)
```

#### TRANSPOSE
And transpose the table above:
```{r}
gdp_2007_t <- data.frame(t(gdp_2007))
knitr::kable(gdp_2007_t)
```

### SORT
Now sort the table "gdp_2007" by "nCountries"

```{r}
gdp_2007_s <- arrange(gdp_2007,nCountries)
knitr::kable(gdp_2007_s)
```



## Join Prompts
### Activty #2 Cheatsheet for dplyr join functions

**Yuki Rang**

#### The data
One data frame is some kpop groups with their group name, number of members, company and representative song.
Another data frame comes from the release year of the songs.
```{r}
kpopGroups <- "
  group,      company, nMembers,repSong
  EXO,         SM,      9,       Growl
  BTS,         BigHit,  7,       Fake Love
  BigBang,      YG,      5,       Fantastic Baby
  GOT7,        JYP,     7,       Hard Carry
  NCT127,      SM,      10,      Boss
  Blackpink,   YG,      4,       Boombahya
  Twice,       JYP,     9,       Cheer Up
  Seventeen,   Pledis,  13,      Don't Wanna Cry
  Red Velvet,  SM,      5,       Red Flavor 
  Wanna One,   YMC,     11,      Energetic
"
kpopGroups <- read_csv(kpopGroups, skip = 1)

releaseYear <- "
  repSong, group, release
  Gee,       Girls's Generation,     2009
  Fantastic Baby, BigBang,  2012
  Growl,      EXO,    2013
  Heart Attack, AOA,  2015
  Hard Carry,  GOT7,   2016
  Boss,        NCT127,   2017
  Fake Love, BTS,     2018
"
releaseYear <- read_csv(releaseYear, skip = 1)
knitr::kable(kpopGroups)
knitr::kable(releaseYear)
```

#### inner_join(kpopGroups,releaseYear)

> 
return all rows from x where there are matching values in y, and all columns from x and y. If there are multiple matches between x and y, all combination of the matches are returned.

```{r}
inner_join(kpopGroups,releaseYear,by='group') %>% 
  knitr::kable()
```

Now we only have five groups listed, because *Wanna One* and so on don't show in the "releaseYear" table and *AOA* and *Girls's Generation* don't show in the "kpopGroups" table. While inner_join() would replicate the repSong column.

#### semi_join(kpopGroups,releaseYear)
> 
return all rows from x where there are matching values in y, keeping just columns from x. A semi join differs from an inner join because an inner join will return one row of x for each matching row of y, where a semi join will never duplicate rows of x.

```{r}
semi_join(kpopGroups,releaseYear,by='group') %>% 
  knitr::kable()
```

We still have five groups listed, while only the repSong from "kpopGroups" has been kept.

#### left_join(kpopGroups,releaseYear)
>
left_join(x, y): Return all rows from x, and all columns from x and y. If there are multiple matches between x and y, all combination of the matches are returned. This is a mutating join.

```{r}
left_join(kpopGroups,releaseYear, by="group") %>% 
  knitr::kable()
```

This time, we take all records from "kpopGroups"", and take the records in "releaseYear" with same group as in "kpopGroups", and combine them together.

Similarly, the right_join()

>return all rows from y, and all columns from x and y. Rows in y with no match in x will have NA values in the new columns. If there are multiple matches between x and y, all combinations of the matches are returned.

```{r}
right_join(kpopGroups,releaseYear, by="group") %>% 
  knitr::kable()
```


#### anti_join(kpopGroups,releaseYear)
>
anti_join(x, y): Return all rows from x where there are not matching values in y, keeping just columns from x. This is a filtering join.

```{r}
anti_join(kpopGroups,releaseYear,by="group") %>% 
  knitr::kable()
```

We keep only five groups in "kpopGroups" that don't show in "releaseYear".

#### full_join(kpopGroups,releaseYear)

> full_join(x, y): Return all rows and all columns from both x and y. Where there are not matching values, returns NA for the one missing. This is a mutating join.

```{r}
full_join(kpopGroups,releaseYear,by="group") %>% 
  knitr::kable()
```

Now we have a full record of all the records from both groups.
